Heliconia IPM Practice

### 2020-09-03 notes/IPM practice.Rmd
moved notebooks to notes directory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)
library(skimr)
library(performance)
library(lme4)
library(bbmle)
library(broom)
library(broom.mixed)
library(popbio)

@@

```{r data, echo=TRUE}
demog <- read_csv(here("analysis", "data", "raw_data", "Ha_survey_with_Zombies.csv"), col_names = TRUE,
               cols(plot = col_character(),
                    bdffp_reserve_no = col_character(),
                    shts = col_integer(),
                    year = col_integer(),
                    infl = col_integer(),
                    tag_number = col_character()))

unique(demog$ranch)
unique(demog$code.notes)
unique(demog$code2)
unique(demog$size)

@@

- `plot` (character):
- `size` (character): fragment size
- `ranch` (character): "Dimona", "Esteio-Colosso", or "PortoAlegre".  The three ranches that make up BDFFP.
- `bdffp_reserve_no` (character): Official BDFFP reserve designation
- `tag_number` (numeric): plant tag# within plot (not unique)
- `row` (character): spatial location in plot
- `column` (numeric): spatial location in plot
- `year` (numeric): year of survey
- `ht` (numeric): plant height.  Plants not straightened before measurement.
- `shts` (numeric): number of shoots
- `infl` (numeric): number of inflorescences
- `code.notes` (character): 
    - "plant missing" = plant was not found in that survey.  May not be dead, may be found in subsequent surveys.
    - "dead" = plant lost all above-ground tissue.  May resprout in a subsequent year.
    - "ULY" = 
    - "sdlg" = seedling, defined as a year 1 plant.
    - "tag missing" = ID tag labeling the plant was missing.  Plants were given new tags and researchers did their best to match up new tag and old tag. **(that column isn't in this version though)**
- `code2` (character): "initial.tag.yr" = marks the first year plant appeared in a survey.  Not necessarily a seedling.

@@

```{r}
km41 %>% group_by(tag_number) %>% summarize(max(year))
max(km41$year)
```

```{r}
km41_2 <- 
  km41 %>%
  group_by(tag_number) %>% 
  # Routine for figuring out last year of data collected
  mutate(x = cumsum(ht > 0 & !is.na(ht))) %>%
  # mutate(x = cumsum(lag(ht) > 0 & !is.na(lag(ht)))) %>%
  mutate(n = 1:n()) %>% 
  mutate(fate = if_else(
    n == which.max(x==max(x)),
    "died",
    NA_character_
  )) %>% 
  # remove rows after plant is dead
  filter(row_number() <= which(fate == "died")[1]) %>% 
  # remove "died" if last year of data (2009)
  ungroup() %>% 
  mutate(fate = if_else(year == max(year), NA_character_, fate)) %>% 
  select(-n, -x)
# km41_2 %>% filter(fate == "died") %>% count(year)
```
```{r}
km41_2 %>% count(fate)
```
359 plants died

@@

```{r}
km41_2 <-
  km41_2 %>% 
  group_by(tag_number) %>% 
  mutate(ht_prev = lag(ht))
```

@@

- each inflorescence has 20--25 flowers
- proportion of flowers developing into fruits = 0.30±0.24 (range = 0--0.88)
- Number of seeds per plant = 4.74±6.10 (range = 0--27)
- Number of seeds per flower = 0.27±0.33 (range = 0--1)

@@

```{r}
meo_pflwr2 <- glmer(flowered ~ ht + I(ht^2) + (1|tag_number) + (1|year), family = binomial, data = km41_2)
```
Check random effects

```{r}
mo_pflwr2 <- glmer(flowered ~ ht + I(ht^2) + (1|tag_number), family = binomial, data = km41_2)
me_pflwr2 <- glmer(flowered ~ ht + I(ht^2) + (1|year), family = binomial, data = km41_2)
m_pflwr2  <-   glm(flowered ~ ht + I(ht^2), family = binomial, data = km41_2)
AICtab(meo_pflwr2, me_pflwr2, mo_pflwr2, m_pflwr2)
```
Keep both random effects.

@@

```{r}
meo_pflwr <- glmer(flowered ~ ht + (1|tag_number) + (1|year), family = binomial, data = km41_2)
AICtab(meo_pflwr, meo_pflwr2)
```

@@

```{r}
meo_flwrs2 <- glmer(infl ~ ht + I(ht^2) + (1|tag_number) + (1|year), family = poisson, data = km41_flowered)
```
```{r}
me_flwrs2 <- glmer(infl ~ ht + I(ht^2) + (1|year), family = poisson, data = km41_flowered)
mo_flwrs2 <- glmer(infl ~ ht + I(ht^2) + (1|tag_number), family = poisson, data = km41_flowered)
m_flwrs2 <- glm(infl ~ ht + I(ht^2), family = poisson, data = km41_flowered)
AICtab(meo_flwrs2, me_flwrs2, mo_flwrs2, m_flwrs2)
```

@@

```{r}
km41_rec_sum <- 
  km41 %>%
  group_by(year) %>% 
  summarize(recruits = sum(code.notes == "sdlg (1)", na.rm = TRUE),
            seeds_est = sum(lag(infl), na.rm = TRUE)) %>% 
  filter(year > 1998)
km41_rec_sum
```

@@

```{r}
km41_rec <- km41 %>% filter(code.notes == "sdlg (1)")
m_rec <- lm(ht ~ 1, data = km41_rec)
```

@@

Try modeling growth as height_t+1 ~ height_t.  Needs a random effect of tag_number, I think?
```{r}
meo_grow2 <- lmer(ht ~ ht_prev + I(ht_prev^2) + (1|tag_number) + (1|year), data = km41_2)
```
Test random effects

```{r}
me_grow2 <- lmer(ht ~ ht_prev + I(ht_prev^2) + (1|year), data = km41_2)
mo_grow2 <- lmer(ht ~ ht_prev + I(ht_prev^2) + (1|tag_number), data = km41_2)
m_grow2 <- lm(ht ~ ht_prev + I(ht_prev^2), data = km41_2)
AICtab(meo_grow2, me_grow2, mo_grow2, m_grow2)
```

@@

### Model
```{r}
meo_surv2 <- glmer(surv ~ ht + I(ht^2) + (1|tag_number) + (1|year), family = binomial, data = km41_2)
```
```{r}
me_surv2 <- glmer(surv ~ ht + I(ht^2) + (1|year), family = binomial, data = km41_2)
mo_surv2 <- glmer(surv ~ ht + I(ht^2) + (1|tag_number), family = binomial, data = km41_2)
m_surv2 <- glm(surv ~ ht + I(ht^2), family = binomial, data = km41_2)
AICtab(meo_surv2, me_surv2, mo_surv2, m_surv2)
```

@@

```{r}
mk_K <- function(m, L, U) {

	# mesh points 
	h <- (U - L)/m
	meshpts <- L + ((1:m) - 1/2) * h
	P <- h * (outer(meshpts, meshpts, P_z1z))
	F <- h * (outer(meshpts, meshpts, F_z1z))
	K <- P + F
	return(list(K = K, meshpts = meshpts, P = P, F = F))
}
```

@@

rownames(sens) <- out$meshpts
colnames(sens) <- out$meshpts
matplot2(sens)
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(conflicted)
library(skimr)
library(performance)
library(lme4)
library(bbmle)
library(broom)
library(broom.mixed)
library(popbio)
library(janitor)

@@

```{r data, echo=TRUE}
demog <- read_csv(here("analysis", "data", "raw_data", "Ha_survey_with_Zombies.csv"), col_names = TRUE,
               cols(plot = col_character(),
                    bdffp_reserve_no = col_character(),
                    shts = col_integer(),
                    year = col_integer(),
                    infl = col_integer(),
                    tag_number = col_character(),
                    HA_ID_Number = col_character())) %>% 
  clean_names()

unique(demog$ranch)
unique(demog$code_notes)
unique(demog$code2)
unique(demog$habitat)

@@

- `plot` (character):
- `habitat` (character): fragment size
- `ranch` (character): "DIM" = Dimona, "EST" = Esteio-Colosso, or "PAL" = PortoAlegre.  The three ranches that make up BDFFP.
- `bdffp_reserve_no` (character): Official BDFFP reserve designation
- `ha_id_number` (numeric): A unique ID number for each plant after resolving duplicate tags and such.
- `tag_number` (numeric): plant tag# within plot (not unique)
- `row` (character): spatial location in plot
- `column` (numeric): spatial location in plot
- `year` (numeric): year of survey
- `ht` (numeric): plant height.  Plants not straightened before measurement.
- `shts` (numeric): number of shoots
- `infl` (numeric): number of inflorescences
- `code.notes` (character): 
    - "plant missing" = plant was not found in that survey.  May not be dead, may be found in subsequent surveys.
    - "dead" = plant lost all above-ground tissue.  May resprout in a subsequent year.
    - "ULY" = 
    - "sdlg" = seedling, defined as a year 1 plant.
    - "tag missing" = ID tag labeling the plant was missing.  Plants were given new tags and researchers did their best to match up new tag and old tag.
- `code2` (character): "initial.tag.yr" = marks the first year plant appeared in a survey.  Not necessarily a seedling.

@@

```{r}
km41 %>% group_by(ha_id_number) %>% summarize(max(year))
max(km41$year)
```

```{r}
km41_2 <- 
  km41 %>%
  group_by(ha_id_number) %>% 
  # Routine for figuring out last year of data collected
  mutate(x = cumsum(ht > 0 & !is.na(ht))) %>%
  # mutate(x = cumsum(lag(ht) > 0 & !is.na(lag(ht)))) %>%
  mutate(n = 1:n()) %>% 
  mutate(fate = if_else(
    n == which.max(x==max(x)),
    "died",
    NA_character_
  )) %>% 
  # remove rows after plant is dead
  filter(row_number() <= which(fate == "died")[1]) %>% 
  # remove "died" if last year of data (2009)
  ungroup() %>% 
  mutate(fate = if_else(year == max(year), NA_character_, fate)) %>% 
  select(-n, -x)
# km41_2 %>% filter(fate == "died") %>% count(year)
```
```{r}
km41_2 %>% count(fate)
```
362 plants died

@@

```{r}
km41_2 <-
  km41_2 %>% 
  group_by(ha_id_number) %>% 
  mutate(ht_prev = lag(ht))
```

@@

- each inflorescence has 20--25 flowers
- proportion of flowers developing into fruits = 0.30±0.24 (range = 0--0.88)
- Number of seeds per plant = 4.74±6.10 (range = 0--27)
- Number of sees per fruit =  1, 2, or 3 with equal probability.
- Number of seeds per flower = 0.27±0.33 (range = 0--1)

@@

```{r}
meo_pflwr2 <- glmer(flowered ~ ht + I(ht^2) + (1|ha_id_number) + (1|year), family = binomial, data = km41_2)
```
Check random effects

```{r}
mo_pflwr2 <- glmer(flowered ~ ht + I(ht^2) + (1|ha_id_number), family = binomial, data = km41_2)
me_pflwr2 <- glmer(flowered ~ ht + I(ht^2) + (1|year), family = binomial, data = km41_2)
m_pflwr2  <-   glm(flowered ~ ht + I(ht^2), family = binomial, data = km41_2)
AICtab(meo_pflwr2, me_pflwr2, mo_pflwr2, m_pflwr2)
```
Keep both random effects.

@@

```{r}
meo_pflwr <- glmer(flowered ~ ht + (1|ha_id_number) + (1|year), family = binomial, data = km41_2)
AICtab(meo_pflwr, meo_pflwr2)
```

@@

```{r}
meo_flwrs2 <- glmer(infl ~ ht + I(ht^2) + (1|ha_id_number) + (1|year), family = poisson, data = km41_flowered)
```
```{r}
me_flwrs2 <- glmer(infl ~ ht + I(ht^2) + (1|year), family = poisson, data = km41_flowered)
mo_flwrs2 <- glmer(infl ~ ht + I(ht^2) + (1|ha_id_number), family = poisson, data = km41_flowered)
m_flwrs2 <- glm(infl ~ ht + I(ht^2), family = poisson, data = km41_flowered)
AICtab(meo_flwrs2, me_flwrs2, mo_flwrs2, m_flwrs2)
```

@@

```{r}
km41_rec_sum <- 
  km41 %>%
  group_by(year) %>% 
  summarize(recruits = sum(code_notes == "sdlg (1)", na.rm = TRUE),
            seeds_est = sum(lag(infl), na.rm = TRUE)) %>% 
  filter(year > 1998)
km41_rec_sum
```

@@

```{r}
km41_rec <- km41 %>% filter(code_notes == "sdlg (1)")
m_rec <- lm(ht ~ 1, data = km41_rec)
```

@@

Try modeling growth as height_t+1 ~ height_t.  Needs a random effect of tag_number, I think?
```{r}
meo_grow2 <- lmer(ht ~ ht_prev + I(ht_prev^2) + (1|ha_id_number) + (1|year), data = km41_2)
```
Test random effects

```{r}
me_grow2 <- lmer(ht ~ ht_prev + I(ht_prev^2) + (1|year), data = km41_2)
mo_grow2 <- lmer(ht ~ ht_prev + I(ht_prev^2) + (1|ha_id_number), data = km41_2)
m_grow2 <- lm(ht ~ ht_prev + I(ht_prev^2), data = km41_2)
AICtab(meo_grow2, me_grow2, mo_grow2, m_grow2)
```

@@

### Model
```{r}
meo_surv2 <- glmer(surv ~ ht + I(ht^2) + (1|ha_id_number) + (1|year), family = binomial, data = km41_2)
```
```{r}
me_surv2 <- glmer(surv ~ ht + I(ht^2) + (1|year), family = binomial, data = km41_2)
mo_surv2 <- glmer(surv ~ ht + I(ht^2) + (1|ha_id_number), family = binomial, data = km41_2)
m_surv2 <- glm(surv ~ ht + I(ht^2), family = binomial, data = km41_2)
AICtab(meo_surv2, me_surv2, mo_surv2, m_surv2)
```

@@

```{r}
mk_K <- function(m, L, U) {
	# mesh points 
	h <- (U - L)/m
	meshpts <- L + ((1:m) - 1/2) * h
	P <- h * (outer(meshpts, meshpts, P_z1z))
	F <- h * (outer(meshpts, meshpts, F_z1z))
	K <- P + F
	return(list(K = K, meshpts = meshpts, P = P, F = F))
}
```

@@

rownames(sens) <- out$meshpts
colnames(sens) <- out$meshpts
matplot2(sens, xlab = "Height")
```

@@

```{r}
elas <- elasticity(out$K)
rownames(elas) <- colnames(elas) <- out$meshpts
matplot2(elas, xlab = "Height")
```
